Bayesian Hierarchical Modeling: An Introduction and Reassessment

### 2022-06-09 F - Tutorial Log-Normal Model brms/Online Supplement F - Bayesian Hierarchical Modeling in brms - Log Model.Rmd
All documents except .rds

---
title             : "Online Supplement D: Bayesian Hierarchical Modeling in brms"
author            : "Myrthe Veenman"
bibliography      : ["r-references.bib"]

@@

|    This online supplement provides a tutorial on Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b]. In addition, it contains information about the Number Classification Task that could not be included in the paper. As `brms` runs on `Stan` [@carpenter2017stan], it is very similar to `rstan` [@R-rstan]. However, there are some differences that will be discussed in this tutorial. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. In concrete, the tutorial will cover the following:

@@

|    To fit a BHM in *brms*, we use the *brm* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = gaussian(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

@@

However, for the digit classification task, we specified the formula in a slightly different way that allowed us to use a different prior on the intercept compared to the distance and side effects. In the first part, we specify that the dependent variable (in our case response times) are predicted by `a` and `b`. In the second part, we define `a`. It contains the general and individual distance and side effects. Also note that we replaced `1` by `0`, indicating that we do not want the default intercept. In the third and final part, we specify `b`, representing the general and individual intercepts by the variable `inter`.

\scriptsize
```{r formulafit2, eval = FALSE}
formula <- brms::bf(rt2 ~ a + b, nl = TRUE,  # part 1 
                  lf(a ~ 0 + side + dif1 + dif2 + dif3 + dif4 + (0 + side + dif1 + dif2 + dif3 + dif4 || ind), center = TRUE),   # part 2 
                  lf(b ~ 0 + inter + (0 + inter || ind), cmc = FALSE)  # part 3 
)
```
\normalsize

@@

|    The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed. For the normal model, this argument takes the value `gaussian`.

@@

|    Finally, we specify the priors. In `brms`, priors can be set on the general effects represented by `class = b` and on the individual variation represented by `class = sd`. It is important to note that priors can only be set on the specific parameters that are part of the brms model specification. For example, `brms` only allows setting priors on standard deviations and not on variances.
Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation 7. To set the priors, we use the `set_prior` function, as shown below. In this function, first, the distribution is specified, using the `Stan` programming language. Then, the type of parameter to which the prior applies is defined by `class`. The parameter to which this prior applies is specified by `coef`. For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., `group`). For example, the prior distributions on the general side effect, $\mu_{\beta}$, and on its variability between individuals, $\sigma_{\beta}$, can be specified as follows: 
&nbsp;

\scriptsize 
```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
set_prior("normal(0,0.3)", class = "b", coef="side")  # general effect 
set_prior("inv_gamma(13.8, 5.3)", class = "sd",       # individual variation 
          coef="side", group = "ind") 
```
\normalsize

For the digit classification task, we want to specify a truncated normal prior on the intercept, because response times cannot be negative. This means that we are setting a different prior on the intercept than on the side and digit effects. To accomplish this, we split up the formula in two different parts. For each variable, we specify to which part of the formula they belong with `nlpar`. In addition, for the intercept, we specify that the prior has a lower bound of zero with `lb`. All priors are saved in an R `data.frame` that is used in the `brm` function.

\scriptsize 
```{r priors}
priorsmodel <- c(set_prior("normal(0.5,1)", class = "b", lb = 0, nlpar = "b"),  
                  set_prior("inv_gamma(13.8, 6.3)", class = "sd", coef="inter", group = "ind", nlpar = "b"), 
                  set_prior("normal(0,0.3)", class = "b", coef="side",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="side", group = "ind",  nlpar = "a"), 
                  set_prior("normal(0,0.3)", class = "b", coef="dif1",  nlpar = "a"), 
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif1", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif2",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif2", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif3",  nlpar = "a"), 
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif3", group = "ind",  nlpar = "a"),
                  set_prior("normal(0,0.3)", class = "b", coef="dif4",  nlpar = "a"),  
                  set_prior("inv_gamma(13.8,5.3)", class = "sd", coef="dif4", group = "ind",  nlpar = "a"),
                  set_prior("inv_gamma(13.8, 6.3)", class = "sigma"))

@@

datalistpp3 <- list()
for (i in 1:1000){
# Hyperpriors: These are not individual specific so only sample 1  
mugammapp <- rtruncnorm(1, a = 0, mean = 0.5, sd = sqrt(1))
vargammapp <- rinvgamma(1, 3, 0.7)
mubetapp <- rnorm(1, 0, 0.3)
varbetapp <- rinvgamma(1, 3, 0.5)
mudeltapp <- rnorm(1, 0, 0.3)
vardeltapp <- rinvgamma(1, 3, 0.5)

@@

sigmapp <- rinvgamma(1, 3, 0.7)  # Only 1 because not individual specific

# data
Ypp <- rnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
              ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
              wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
              sqrt(sigmapp))  # model formula 
datartpp <- data.frame(exppp,  # data frame of results and conditions 
                       condpp, 
                       subpp, 
                       repetition = rep(i, length(exppp)), 
                       gammaipp[exppp], betaipp[exppp], delta1pp[exppp],
                       delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
datalistpp3[[i]] <- datartpp
}

@@

newbigdatapp3 <- bigdatapp3[bigdatapp3$exppp == 1,]

@@

\scriptsize 
``` {r modelfit, eval = FALSE}
model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = gaussian(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 # sample_prior = "yes",                 # To calculate Bayes factors for point hypotheses  
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

@@

We have already estimated the model and saved it in the R object `brmsmodelfit_final`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
model_fit <- readRDS(file = "myPath/brmsmodelfit_final.rds")

# For model comparison - sample_prior was set to "yes" 
model_fit_bf <- readRDS("myPath/brmsmodelfit_BF_final.rds")

@@

\scriptsize 
```{r loadmodelfits, show = FALSE}
model_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_03012022.rds")

# For model comparison - sample_prior was set to "yes" 
model_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/Truncated/BRMSmodelfit18_BF_03012022.rds")

@@

\scriptsize 
```{r modelsummary}
summary(model_fit)
```
\normalsize

@@

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(model_fit)
```
\normalsize

@@

\scriptsize 
```{r checkvarnames, eval = FALSE}
variables(model_fit)
```
\normalsize

@@

# Default trace plot for parameter mu4
# I think warmup already deleted, but then have to adjust labels x axis 
trdif2brms <- mcmc_plot(model_fit,       # Model fit 
                       variable = "b_a_dif2",  # Select parameter, 
                       # it is also possible to select more than one at the same time, 
                       # such as c("mu2", "mu3","mu4")
                       type = "trace")   # Specify that you want trace plots

# Function 
traceplotfunc <- function(parameter, title){
trdif2brms <- mcmc_plot(model_fit,       # Model fit 
                       variable = parameter,  # Select parameter, 
                       type = "trace")   # Specify that you want trace plots

@@

# Show trace plot for one digit parameter 
tr2 <- traceplotfunc(parameter = "b_a_side", title = expression(~mu[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "b_a_dif1", title = expression(~mu[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "b_a_dif2", title = expression(~mu[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "b_a_dif3", title = expression(~mu[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "b_a_dif4", title = expression(~mu[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

@@

\scriptsize 
```{r traceplotsd, }
# Show trace plot for one sd parameter 
tr2 <- traceplotfunc(parameter = "sd_sub1__a_side", title = expression(~sigma[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "sd_sub1__a_dif1", title = expression(~sigma[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "sd_sub1__a_dif2", title = expression(~sigma[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "sd_sub1__a_dif3", title = expression(~sigma[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "sd_sub1__a_dif4", title = expression(~sigma[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

@@

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
mcmc_plot(model_fit, type = "rhat_hist")   # Specify that you want trace plots

# Plot frequency of the ratio 
mcmc_plot(model_fit, type = "neff_hist") 
```
\normalsize

@@

\scriptsize 
```{r rhatexamp}
# Obtain rhat and neff brms 
## Rhat 
rhatbrms <- mcmc_plot(model_fit, type = "rhat_hist")$data$value  # save rhat values 
rhatbrms_para <- mcmc_plot(model_fit, type = "rhat_hist")$data$parameter  # save parameter names 
rhatbrmsratio_comb <- data.frame(parameter = rhatbrms_para, value = rhatbrms)  # combine values and parameters names into one data.frame 
```
\normalsize

@@

\scriptsize 
```{r rhatexamp2}
# Obtain rhats of general effects 
rhatbrms_v2 <- mcmc_plot(model_fit, type = "rhat_hist")$data
fixedrhatbrms_v2 <- c(rhatbrms_v2$value[rhatbrms_v2$parameter == "b_b_inter"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_side"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif1"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif2"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif3"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_a_dif4"])

@@

\scriptsize 
```{r rhatexamp3}
# Obtain rhats of all individual effects 
rhatbrmsratio3 <- c(rhatbrmsratio_comb[grep(",inter]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",side]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif1]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif2]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif3]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif4]", rhatbrmsratio_comb$parameter),]$value)

@@

\scriptsize 
```{r neffexamp5}
neffbrmsratio_value <- mcmc_plot(model_fit, type = "neff_hist")$data$value   # save neff values, these are the values of the ration Neff/N, N representing the total number of iterations
neffbrmsratio_para <- mcmc_plot(model_fit, type = "neff_hist")$data$parameter  # save parameter names 
neffbrmsratio_comb <- data.frame(parameter = neffbrmsratio_para, value = neffbrmsratio_value)  # combine values and parameter names into data frame

# Neff of general effects 
fixedneffbrms2 <- c(neffbrmsratio_comb[grep("b_inter", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_side", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif1", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif2", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif3", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_a_dif4", neffbrmsratio_comb$parameter),]$value[1])

@@

# Neff of individual effects  
neffbrmsratio3 <- c(neffbrmsratio_comb[grep(",inter]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",side]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif1]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif2]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif3]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif4]", neffbrmsratio_comb$parameter),]$value)
neffbrms2 <- neffbrmsratio3  * 12000  # we multiply values with total number of iterations (12000) to get Neff (used to be ration Neff/N, N representing the total number of iterations)

@@

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
mcmc_plot(model_fit, 
     type = "dens")  
```
\normalsize

@@

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
post8 <- as.array(model_fit)  # change structure of model fit

@@

pairs8 <- mcmc_areas(
  post8,    # Structured model fit 
  pars = c("b_a_side", "b_a_dif1", "b_a_dif2", "b_a_dif3", "b_a_dif4"),  # Select parameters 
  prob = 0.95, # 80% intervals  # Plot 95% credible intervals
  prob_outer = 0.99, # 99%     # 99% outer interval
  point_est = "mean"           # Plot line in the middle 
)
# Returns ggplot object

@@

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(model_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)
fitbrmshiertable2 <- rbind(fitbrmshiertable2[6,], fitbrmshiertable2[1:5,])

@@

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(model_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)
fitbrmshiertable22 <- rbind(fitbrmshiertable22[6,], fitbrmshiertable22[1:5,])

@@

\scriptsize 
```{r randefplotexample}
# Gamma 
# Extract the parameter estimates 
brmsgammar <- coef(model_fit)$sub1[,,"b_inter"][,c("Estimate", "Q2.5", "Q97.5")]
brmsgammar <- as.data.frame(brmsgammar)

# Extract the general effect estimate of the parameter 
fixbremsgammar <- fixef(model_fit)["b_inter","Estimate"]

@@

# Beta 
brmsbetar <- coef(model_fit)$sub1[,,"a_side"][,c("Estimate", "Q2.5", "Q97.5")]
brmsbetar <- as.data.frame(brmsbetar)
fixbremsbetar <- fixef(model_fit)["a_side","Estimate"]

>>>>

---
title             : "Online Supplement F: Bayesian Hierarchical Modeling in brms - Log Model"
author            : "Myrthe Veenman"
bibliography      : ["r-references.bib"]

@@

|    This online supplement provides a tutorial on log-normal Bayesian Hierarchical models (BHM) in `brms` [@R-brms_a; @R-brms_b]. In addition, it contains information about the Number Classification Task that could not be included in the paper. As `brms` runs on `Stan` [@carpenter2017stan], it is very similar to `rstan` [@R-rstan]. However, there are some differences that will be discussed in this tutorial. Only the fit of the full model will be shown as the fitting the others should follow logically from this tutorial. Technical details, such as explanations of statistical terms (e.g., $\hat{R}$), will not be discussed here but can be found in the paper. In concrete, the tutorial will cover the following:

@@

|    To fit a log-normal BHM in *brms*, we use the *brm* function, as shown below. 
&nbsp;

\scriptsize 
```{r modelfitpreview, eval=FALSE, echo=TRUE}
model_fit <- brm(formula = formula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = lognormal(),                    # Family, distribution dependent variable 
                 prior = priorsmodel,                    # Specification of the priors 
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

@@

@@

|    The argument `family` in the `brm` function specifies the exponential family of distributions according to which the dependent variable is assumed to be distributed. For the log-normal model, this argument takes the value `lognormal`.

@@

|    Finally, we specify the priors. In `brms`, priors can be set on the general effects represented by `class = b` and on the individual variation represented by `class = sd`. It is important to note that priors can only be set on the specific parameters that are part of the brms model specification. For example, `brms` only allows setting priors on standard deviations and not on variances.
Therefore, for the normal model of the symbolic distance effect, we will use the priors for standard deviations specified in Equation 9. To set the priors, we use the `set_prior` function, as shown below. In this function, first, the distribution is specified, using the `Stan` programming language. Then, the type of parameter to which the prior applies is defined by `class`. The parameter to which this prior applies is specified by `coef`. For the priors on the individual effects, an additional element has to be specified, namely, the grouping variable (i.e., `group`). For example, the prior distributions on the general side effect, $\mu_{\beta}$, and on its variability between individuals, $\sigma_{\beta}$, can be specified as follows: 
&nbsp;

\scriptsize 
```{r priorsillusbrms2, eval = FALSE, echo = TRUE}
set_prior("normal(-0.5, 1)", class = "b", coef="side")  # general effect 
set_prior("inv_gamma(13.8, 4.1)", class = "sd",       # individual variation 
          coef="side", group = "ind") 
```
\normalsize

All priors are saved in an R `data.frame` that is used in the `brm` function.

\scriptsize 
```{r priors}
priorslogmodel <- c(set_prior("normal(-0.5, 1)", class = "Intercept"),  
                  set_prior("inv_gamma(13.8, 4.1)", class = "sd", coef="Intercept", group = "sub1"), 
                  set_prior("normal(0, 0.07)", class = "b", coef="side"),  
                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="side", group = "sub1"), 
                  set_prior("normal(0, 0.07)", class = "b", coef="dif1"), 
                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif1", group = "sub1"),
                  set_prior("normal(0, 0.07)", class = "b", coef="dif2"),  
                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif2", group = "sub1"),
                  set_prior("normal(0, 0.07)", class = "b", coef="dif3"), 
                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif3", group = "sub1"),
                  set_prior("normal(0, 0.07)", class = "b", coef="dif4"),  
                  set_prior("inv_gamma(13.8, 0.7)", class = "sd", coef="dif4", group = "sub1"),
                  set_prior("inv_gamma(13.8, 4.1)", class = "sigma"))

@@

datalistpp3 <- list()
delta1pp2 <- list()
gammaipp2 <- list()
betaipp2 <- list()
sigmapp2 <- c()
for (i in 1:1000){  # with 10000 my computer cannot handle it 
  # Hyperpriors: These are not individual specific right? So only sample 1? 
  mugammapp <- rnorm(1, -0.5, sqrt(1))
  vargammapp <- rinvgamma(1, 13.8, 4.1)
  mubetapp <- rnorm(1, 0, 0.07)
  varbetapp <- rinvgamma(1, 13.8, 0.7)
  mudeltapp <- rnorm(1, 0, 0.07)
  vardeltapp <- rinvgamma(1, 13.8, 0.7)

@@

  sigmapp <- rinvgamma(1, 13.8, 4.1)  # Only 1 because not individual specific?

  # Or start for loop here? Or do we want to vary persons as well? 
  # data
  Ypp <- rlnorm(Ipp*Jpp*Kpp, gammaipp[exppp] + xjpp * betaipp[exppp] + 
                  ujpp * delta1pp[exppp] + vjpp * delta2pp[exppp] +
                  wjpp * delta3pp[exppp] + zjpp * delta4pp[exppp],
                sqrt(sigmapp))
  gammaipp2[[i]] <- gammaipp
  betaipp2[[i]] <- betaipp
  delta1pp2[[i]] <- delta1pp
  sigmapp2[i] <- sigmapp
  datartpp <- data.frame(exppp, condpp, subpp, repetition = rep(i, length(exppp)), gammaipp[exppp], betaipp[exppp], delta1pp[exppp], delta2pp[exppp], delta3pp[exppp], delta4pp[exppp], Ypp)
  datalistpp3[[i]] <- datartpp
}

@@

@@

\scriptsize 
``` {r modelfit, eval = FALSE}
logmodel_fit <- brm(formula = logformula,                      # Model formula 
                 data = myData_dataframe,                # Dataset (parameters, observations) 
                 family = lognormal(),                    # Family, distribution dependent variable 
                 prior = priorslogmodel,                    # Specification of the priors 
                 # sample_prior = "yes",                 # To calculate Bayes factors for point hypotheses  
                 warmup = 1000,                          # Nr. iterations for warmup (per chain)
                 iter = 4000,                            # Nr. iterations (per chain)
                 chains = 4,                             # Nr. chains 
                 core = 4)                               # Nr. of cores for parallel estimation

@@

We have already estimated the model and saved it in the R object `brmslogmodelfit_final`. In case you do not want to wait until the model has been fitted but continue immediately with the rest of this tutorial, you can load this R object that contains the model estimations.  
&nbsp;

\scriptsize 
```{r loadmodelfit, eval = FALSE}
logmodel_fit <- readRDS(file = "myPath/brmslogmodelfit_final.rds")

# For model comparison - sample_prior was set to "yes" 
logmodel_fit_bf <- readRDS("myPath/brmslogmodelfit_BF_final.rds")

@@

\scriptsize 
```{r loadmodelfits, show = FALSE}
logmodel_fit <- readRDS(file = "/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/non-linear/brmmodel8log13012022.rds")

# For model comparison - sample_prior was set to "yes" 
logmodel_fit_bf <- readRDS("/Users/myrtheveenman/Documents/GitHub/multilevel-tutorial-paper/R objects/brms/BF/brmmodel8log26052022_bf.rds")

@@

\scriptsize 
```{r modelsummary}
summary(logmodel_fit)
```
\normalsize

@@

\scriptsize 
```{r shinyrstan, eval = FALSE}
launch_shinystan(logmodel_fit)
```
\normalsize

@@

\scriptsize 
```{r checkvarnames, eval = FALSE}
variables(logmodel_fit)
```
\normalsize

@@

# Default trace plot for parameter mu4
# I think warmup already deleted, but then have to adjust labels x axis 
trdif2brms <- mcmc_plot(logmodel_fit,       # Model fit 
                       variable = "b_dif2",  # Select parameter, 
                       # it is also possible to select more than one at the same time, 
                       # such as c("mu2", "mu3","mu4")
                       type = "trace")   # Specify that you want trace plots

# Function 
traceplotfunc <- function(parameter, title){
trdif2brms <- mcmc_plot(logmodel_fit,       # Model fit 
                       variable = parameter,  # Select parameter, 
                       type = "trace")   # Specify that you want trace plots

@@

# Show trace plot for one digit parameter 
tr2 <- traceplotfunc(parameter = "b_side", title = expression(~mu[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "b_dif1", title = expression(~mu[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "b_dif2", title = expression(~mu[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "b_dif3", title = expression(~mu[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "b_dif4", title = expression(~mu[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

@@

\scriptsize 
```{r traceplotsd, }
# Show trace plot for one sd parameter 
tr2 <- traceplotfunc(parameter = "sd_sub1__side", title = expression(~sigma[~beta]))
tr2 <- tr2 + theme(legend.position = "none")
tr3 <- traceplotfunc(parameter = "sd_sub1__dif1", title = expression(~sigma[~delta[~7]]))
tr3 <- tr3 + theme(legend.position = "none")
tr4 <- traceplotfunc(parameter = "sd_sub1__dif2", title = expression(~sigma[~delta[~6]]))
tr4 <- tr4 + theme(legend.position = "none")
tr5 <- traceplotfunc(parameter = "sd_sub1__dif3", title = expression(~sigma[~delta[~4]]))
tr5 <- tr5 + theme(legend.position = "none")
tr6 <- traceplotfunc(parameter = "sd_sub1__dif4", title = expression(~sigma[~delta[~3]]))
legendtr <- cowplot::get_legend(tr6)
tr6 <- tr6 + theme(legend.position = "none")

@@

\scriptsize 
```{r rhatneffstandard, eval = FALSE}
# Plot frequency of rhat 
mcmc_plot(logmodel_fit, type = "rhat_hist")   # Specify that you want trace plots

# Plot frequency of the ratio 
mcmc_plot(logmodel_fit, type = "neff_hist") 
```
\normalsize

@@

\scriptsize 
```{r rhatexamp}
# Obtain rhat and neff brms 
## Rhat 
rhatbrms <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data$value  # save rhat values 
rhatbrms_para <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data$parameter  # save parameter names 
rhatbrmsratio_comb <- data.frame(parameter = rhatbrms_para, value = rhatbrms)  # combine values and parameters names into one data.frame 
```
\normalsize

@@

\scriptsize 
```{r rhatexamp2}
# Obtain rhats of general effects 
rhatbrms_v2 <- mcmc_plot(logmodel_fit, type = "rhat_hist")$data
fixedrhatbrms_v2 <- c(rhatbrms_v2$value[rhatbrms_v2$parameter == "b_Intercept"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_side"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif1"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif2"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif3"], 
                      rhatbrms_v2$value[rhatbrms_v2$parameter == "b_dif4"])

@@

\scriptsize 
```{r rhatexamp3}
# Obtain rhats of all individual effects 
rhatbrmsratio3 <- c(rhatbrmsratio_comb[grep(",Intercept]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",side]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif1]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif2]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif3]", rhatbrmsratio_comb$parameter),]$value,
                      rhatbrmsratio_comb[grep(",dif4]", rhatbrmsratio_comb$parameter),]$value)

@@

\scriptsize 
```{r neffexamp5}
neffbrmsratio_value <- mcmc_plot(logmodel_fit, type = "neff_hist")$data$value   # save neff values, these are the values of the ration Neff/N, N representing the total number of iterations
neffbrmsratio_para <- mcmc_plot(logmodel_fit, type = "neff_hist")$data$parameter  # save parameter names 
neffbrmsratio_comb <- data.frame(parameter = neffbrmsratio_para, value = neffbrmsratio_value)  # combine values and parameter names into data frame

# Neff of general effects 
fixedneffbrms2 <- c(neffbrmsratio_comb[grep("b_Intercept", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_side", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_dif1", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_dif2", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_dif3", neffbrmsratio_comb$parameter),]$value[1],
                      neffbrmsratio_comb[grep("b_dif4", neffbrmsratio_comb$parameter),]$value[1])

@@

# Neff of individual effects  
neffbrmsratio3 <- c(neffbrmsratio_comb[grep(",Intercept]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",side]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif1]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif2]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif3]", neffbrmsratio_comb$parameter),]$value,
                      neffbrmsratio_comb[grep(",dif4]", neffbrmsratio_comb$parameter),]$value)
neffbrms2 <- neffbrmsratio3  * 12000  # we multiply values with total number of iterations (12000) to get Neff (used to be ration Neff/N, N representing the total number of iterations)

@@

\scriptsize 
```{r posteriorstandard, eval = FALSE}
# Seperate plot per parameter of posterior density 
mcmc_plot(logmodel_fit, 
     type = "dens")  
```
\normalsize

@@

\scriptsize 
```{r posteriormanual, fig.cap="The posterior distributions for the general effects. The middle line within the distributions represents the mean posterior. The shaded area within the distributions represents 95\\% of the probability mass.", fig.align = "left", fig.height = 4, fig.width = 5}
post8 <- as.array(logmodel_fit)  # change structure of model fit

@@

pairs8 <- mcmc_areas(
  post8,    # Structured model fit 
  pars = c("b_side", "b_dif1", "b_dif2", "b_dif3", "b_dif4"),  # Select parameters 
  prob = 0.95, # 80% intervals  # Plot 95% credible intervals
  prob_outer = 0.99, # 99%     # 99% outer interval
  point_est = "mean"           # Plot line in the middle 
)
# Returns ggplot object

@@

\scriptsize 
```{r geneftable, results = "asis", cache = TRUE}
# Obtain the mean, standard error, 95% credible interval, rhat and number of effective samples for the intercept, and the digit and side parameters 
fitbrmshiertable <- summary(logmodel_fit)[[14]][,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]  # variance

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable2 <- as.data.frame(fitbrmshiertable)

@@

\scriptsize 
```{r inddevtable, results = "asis", cache = TRUE}
# Obtain standard deviation estimates 
fitbrmshiertable2 <- summary(logmodel_fit)[[17]]$sub1[,c("Estimate", "Est.Error", "l-95% CI", "u-95% CI", "Bulk_ESS", "Rhat")]

# The results are transformed to a dataframe that allows us to plot the results nicely in a table 
fitbrmshiertable22 <- as.data.frame(fitbrmshiertable2)

@@

\scriptsize 
```{r randefplotexample}
# Gamma 
# Extract the parameter estimates 
brmsgammar <- coef(logmodel_fit)$sub1[,,"Intercept"][,c("Estimate", "Q2.5", "Q97.5")]
brmsgammar <- as.data.frame(brmsgammar)

# Extract the general effect estimate of the parameter 
fixbremsgammar <- fixef(logmodel_fit)["Intercept","Estimate"]

@@

# Beta 
brmsbetar <- coef(logmodel_fit)$sub1[,,"side"][,c("Estimate", "Q2.5", "Q97.5")]
brmsbetar <- as.data.frame(brmsbetar)
fixbremsbetar <- fixef(logmodel_fit)["side","Estimate"]
